  
---
title: "What predict the best of the specified execution of the exercise?"
output: html_document
---
##Step 1.1: organize the data: removing the NAs
  we check on summary(pml_train) and locate all the variabls with huge numbers of NAs, and remove them.
```{r,results='hide'}
pml_train <- read.csv("pml-training.csv")
pml_test <- read.csv("pml-testing.csv")
summary(pml_train)
vars.for.removed = c(17:36,50:59,75:83,93:94,96:97,99:101,103:112,127,131:132,134:135,137:138,141:150)
pml_trainNARM <- pml_train[,-vars.for.removed]
pml_testNARM <- pml_test[,-vars.for.removed]
```

##Step 1.2: organize the data: 
  then use describe() to locate variables with very low sd of even equal to zero for the whole set of data
```{r,results='hide',message=FALSE}
library(psych)
des <- data.frame(describe(pml_trainNARM))
vars.for.removed.sd2=c(2:6,12:16,39:44,48:55,69:76)
pml_trainRM2 <- pml_trainNARM[,-vars.for.removed.sd2]
pml_testRM2 <- pml_testNARM[,-vars.for.removed.sd2]
```


## step 2: split the pml_train 
   to train and test data sets for further analysis
```{r,results='hide',message=FALSE}
library(caret)
inTrain <- createDataPartition(y=pml_trainRM2$classe,
                              p=0.7, list=FALSE)
training <- pml_trainRM2[inTrain,]
testing <- pml_trainRM2[-inTrain,]
```
##step 3: fit the model with rpart with all the left variables
```{r,message=FALSE}
library(e1071)
modFit <- train(classe ~ .,method="rpart",data=training)
print(modFit$finalModel)
```


##step 4: draw the plot of the model we fit
  the plot shows X being the most important predictor
```{r,eval=FALSE,message=FALSE}
library(rattle)
fancyRpartPlot(modFit$finalModel)
```

##step 5: predict the test set (the one splited from pml_train)
  By using rpart, we get an accuracy rate of 0.6616822
```{r}
modFitPred <-predict(modFit,newdata=testing)
table(modFitPred,testing$classe)
accucacy= (1674+1138+1082)/nrow(testing)
```

```{r}
accucacy
modFitPredsmall <-predict(modFit,newdata=pml_testRM2 )
```


##step 6.1:  Fit the model according the the cv trained
  we have 0.22 largest cp with smallest variance
```{r,message=FALSE}
library(caret)
set.seed(111)
fitControl = trainControl( method = "cv", number = 10 )
cartGrid = expand.grid( .cp = (1:50)*0.01) 
train(classe ~ . , data = training, method = "rpart",
trControl = fitControl, tuneGrid = cartGrid )
exerciseTREECV = rpart(classe ~ ., data = training, 
control=rpart.control(cp = 0.22))
```

##step 6.2: predict the accuracy using model trained with cv
  we get an accuracy of 0.9994902, much higher than the previous raw model.
```{r}
modFitPred <-predict(exerciseTREECV,newdata=testing,type ="class")
table(modFitPred,testing$classe)
accucacy= (1674+1138+1026+962+1082)/nrow(testing) 
```
```{r}
accucacy 
modFitPredsmall <-predict(modFit,newdata=pml_testRM2 )
```



##step 7.1: fit randomForest model 
  From the importance function, we conclude that "X" is the most important variable in deciding the correctness of the posture, second by num_window,roll_belt,yaw_belt and pitch_belt. This model make a prediction as good as the cv model. 
```{r,results='hide',message=FALSE}
library(randomForest)
exerciseForest = randomForest(classe ~ ., data = training,imiportance=TRUE,na.action=na.omit,mtry=3)
round(importance(exerciseForest),2)
```

```{r}
round(importance(exerciseForest),2)
```


##step 7.2 predict the testing dataset outcome
```{r}
PredictForest = predict(exerciseForest, newdata = testing)
table(testing$classe, PredictForest)
accuracy= (1674+1138+1025+963+1082)/nrow(testing)
```
```{r}
accuracy
PredictForest = predict(exerciseForest, newdata = pml_test)
```




##Fig.1. prp plot of general rpart model
```{r,echo=FALSE,message=FALSE}
library(rattle)
fancyRpartPlot(modFit$finalModel)
```

 

 
##Fig.2. prp plot of the cv model
```{r,echo=FALSE,message=FALSE}
prp(exerciseTREECV)
```
